---
title: "PostQC.Normalization.Processing"
author: "Angie Boysen"
date: "October 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(tidyr)
require(graphics); require(grDevices)
library(Hmisc)
library(gtools)
library(cowplot)
require(RColorBrewer)
library(xlsx)
library(readr)
library(plotly)
library(stringr)
library(GGally)
library(dplyr)
```

## exploring the cyano diel 2 data

#Import data - set filename with QC output here!
```{r}
filename <- "QC_output_wBlkSubDiel_both_Cyano_DCM_all.csv"
mydata.overloaded <- read_csv(filename,comment = "#")
mydata <-mydata.overloaded[,2:12]
glimpse(mydata)
```

##Read in Internal Standard list
Change the master list if you need to!
Also be sure to adjust the numbers in the dine defining IS.aq.list
so that you actually use the internal standards that are added for
your relevant data.
```{r}
IS.cat <- read_csv("../CYANO_MasterList_Summer2016.csv")  
IS.cat <- IS.cat %>% filter(Group == "Internal Std") %>% 
     select(`Compound Name`) %>% 
     unique()
IS.list <- IS.cat[[1]]
# IS.aq.list <- IS.cat[[1]][c(1:4,6)] ## Good for cyano aq
# IS.aq.inj.only <- IS.cat[[1]][c(3,6)] ## Good for cyano aq

IS.aq.list <- IS.cat[[1]][c(3,6,8)] ## Good for cyano dcm
IS.aq.inj.only <- IS.cat[[1]][c(3,6)] ## Good for cyano dcm
```

##look at extraction replication
if things don't look good with some IS make sure to fix them
or at least not include them as possibilities for normalization
```{r, echo=FALSE}
IS.dat <- mydata[grepl(paste(IS.aq.list, collapse="|"), mydata$Compound.Name),]
IS.dat <- IS.dat %>% select(Compound.Name, Area, Replicate.Name) %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     filter(!grepl("Poo",Replicate.Name)) 
ggplot(IS.dat, aes(x=Replicate.Name, y=Area)) + geom_bar(stat="identity") + facet_wrap( ~Compound.Name, scales="free_y")+theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5), legend.position = "top")  + ggtitle("IS Raw Areas")
```

##Edit data 
26R2, 27R1, 27R2, 27R3, 28R1, 28R3, 29R2, 30R1, 34R1, 34R3,
FilterBlk_7-18 
all had a total of 60 rather than 40 uL of extraction stds added
And the vials had a final volum of 421.05 uL,
so the peaks need to be divided by 1.425 for arachidonic acid.
injection stds should be at the right concentrations.

Then, 28R3, 29R2, 32R3, 35R1, 35R2 have their extraction stds
cut by 1.5 again! 1.5 here seems approximately right but it 
is hard to say. this is based on observations across the three
extraction stds rather than notes from the extraction log.

NOTES from AQ [Then, the injection standards for 26R3 are super low (maybe about 1/3 of what they should be), and 28R2, 29R3, and 31R3 are all low (maybe half or 2/3 of what they should be) -din't fix this]

Extraction Standards are doubled in 37R1 and  39R1

Injection Stds are doubled in 26R2

And so 5R1 has half the D8-arachidonic acid
```{r correctISErrors}
aq.data.fixed <- mydata
errors <- mydata %>%
     filter(Replicate.Name=="160721_Smp_26R2_DCM"|
                 Replicate.Name=="160721_Smp_27R1_DCM"|
                 Replicate.Name=="160721_Smp_27R2_DCM"|
                 Replicate.Name=="160721_Smp_27R3_DCM"|
                 Replicate.Name=="160721_Smp_28R1_DCM"|
                 Replicate.Name=="160721_Smp_28R3_DCM"|
                 Replicate.Name=="160721_Smp_29R2_DCM"|
                 Replicate.Name=="160721_Smp_30R1_DCM"|
                 Replicate.Name=="160721_Smp_34R1_DCM"|
                 Replicate.Name=="160721_Smp_34R3_DCM"|
                 Replicate.Name=="160721_Blk_7-18_FilterBlk_DCM") %>% 
     filter(Compound.Name =="D8-Arachidonic Acid") %>%
     mutate(Area = Area/1.425)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name=="160721_Smp_26R2_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_27R1_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_27R2_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_27R3_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_28R1_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_28R3_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_29R2_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_30R1_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_34R1_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_34R3_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Blk_7-18_FilterBlk_DCM") &
                  (aq.data.fixed$Compound.Name==
                        "D8-Arachidonic Acid")
             ] <- errors$Area

errors <- aq.data.fixed %>%
     filter(Replicate.Name=="160721_Smp_28R3_DCM"|
                 Replicate.Name=="160721_Smp_29R2_DCM"|
                 Replicate.Name=="160721_Smp_32R3_DCM"|
                 Replicate.Name=="160721_Smp_35R1_DCM"|
                 Replicate.Name=="160721_Smp_35R2_DCM") %>% 
     filter(Compound.Name=="D8-Arachidonic Acid") %>%
     mutate(Area = Area/1.5)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name=="160721_Smp_28R3_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_29R2_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_32R3_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_35R1_DCM"|
                 aq.data.fixed$Replicate.Name=="160721_Smp_35R2_DCM") &
                  (aq.data.fixed$Compound.Name==
                        "D8-Arachidonic Acid")
             ]  <- errors$Area


errors <- aq.data.fixed %>%
     filter(Replicate.Name=="160728_Smp_37R1_DCM"|
                 Replicate.Name=="160728_Smp_39R1_DCM") %>% 
     filter(Compound.Name=="D8-Arachidonic Acid") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name=="160728_Smp_37R1_DCM"|
                 aq.data.fixed$Replicate.Name=="160728_Smp_39R1_DCM") &
                  (aq.data.fixed$Compound.Name==
                        "D8-Arachidonic Acid")
             ]  <- errors$Area

errors <- aq.data.fixed %>%
     filter(Replicate.Name=="160721_Smp_26R2_DCM") %>% 
     filter(Compound.Name=="D-Phenylalanine"|
                 Compound.Name=="Vitamin B2_IS") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name=="160721_Smp_26R2_DCM") &
                  (aq.data.fixed$Compound.Name==
                        "D-Phenylalanine"|
                 aq.data.fixed$Compound.Name =="Vitamin B2_IS")
             ] <- errors$Area


errors <- aq.data.fixed %>%
     filter(Replicate.Name=="160701_Smp_5R1_DCM") %>% 
     filter(Compound.Name=="D8-Arachidonic Acid") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name=="160701_Smp_5R1_DCM") &
                  (aq.data.fixed$Compound.Name==
                        "D8-Arachidonic Acid")
             ]  <- errors$Area

```

##Rename Runs
Name structure must be:
Date_type_ID_replicate
```{r echo=FALSE}
aq.data.fixed <- aq.data.fixed %>% 
     mutate(Replicate.Name = Replicate.Name %>%
                 str_replace("_DCM",
                             "") %>%
                 str_replace("R1",
                             "_1") %>%
                 str_replace("R2",
                             "_2") %>%
                 str_replace("R3",
                             "_3") %>%
                 str_replace("160721_Blk_4",
                             "160721_Blk_Blk_4") %>%
                 str_replace("160721_Blk_18",
                             "160721_Blk_Blk_18") %>%
                 str_replace("160721_Blk_7-19_FilterBlk",
                             "160721_Blk_FilterBlk_7-19") %>%
                 str_replace("160721_Blk_7-18_FilterBlk",
                             "160721_Blk_FilterBlk_7-18") %>%
                 str_replace("160728_Blk_7-26_FilterBlk",
                             "160728_Blk_FilterBlk_7-26") %>%
                 str_replace("160728_Blk_18",
                             "160728_Blk_Blk_18")%>%
                 str_replace("160728_Blk_24",
                             "160728_Blk_Blk_24")%>%
                 str_replace("160721_Poo_AllCyanoDCM-Extracts",
                             "160721_Poo_AllCyanoDCM-Extracts4")%>%
                 str_replace("160728_Poo_AllCyanoDCM-Extracts",
                             "160728_Poo_AllCyanoDCM-Extracts5")%>%
                 str_replace("160701_Poo_AllCyanoDCMExtracts",
                             "160701_Poo_AllCyanoDCM-Extracts1")%>%
                 str_replace("160708_Poo_AllCyanoDCMExtracts",
                             "160708_Poo_AllCyanoDCM-Extracts2")%>%
                 str_replace("160714_Poo_AllCyanoDCM-Extracts",
                             "160714_Poo_AllCyanoDCM-Extracts3")%>%
                 str_replace("160714_Blk_6",
                             "160714_Blk_Blk_6") %>%
                 str_replace("160714_Blk_13",
                             "160714_Blk_Blk_13") %>%
                 str_replace("160714_Blk_7-11_FilterBlk",
                             "160714_Blk_FilterBlk_7-11")%>%
                 str_replace("160708_Blk_16",
                             "160708_Blk_Blk_16") %>%
                 str_replace("160708_Blk_4",
                             "160708_Blk_Blk_4") %>%
                 str_replace("160701_Blk_5",
                             "160701_Blk_Blk_5") %>%
                 str_replace("160701_Blk_16",
                             "160701_Blk_Blk_16"))
```

#Break Up the Names
Name structure must be:
Date_type_ID_replicate
```{r echo=FALSE}
mydata_new <- aq.data.fixed %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type", "SampID",
                                        "replicate"),"_") %>%
     mutate(Run.Cmpd = paste(aq.data.fixed$Replicate.Name, 
                             aq.data.fixed$Compound.Name),
            Replicate.Name = aq.data.fixed$Replicate.Name)
mydata <- mydata %>% mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))
aq.data.fixed <- aq.data.fixed %>% mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))
```

#remove negative values from the blank subtracted data
```{r aq.noNegs}
mydata_new <- mydata_new %>% 
     filter(Area.BlkSub>500 | 
                 is.na(Area.BlkSub) |
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(BlkRatio<0.5 |
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("S/N",Notes)|
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>% 
     filter(!grepl("too small",Notes)|
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("No second trace",Notes) |
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("Bad RT",Notes))
```

##Pull out compounds that were detected
```{r aq.detected?, echo=F}
num <- length(unique(mydata_new$Replicate.Name))-6
aq.detect <- mydata_new %>%      
     filter(!grepl("Blk",Replicate.Name)) %>%
     select(Replicate.Name, Compound.Name, Area.BlkSub, Area) %>%
     group_by(Compound.Name) %>%
     summarise(percent.dcm.blksub = sum(!is.na(Area.BlkSub))/num,
               percent.dcm.raw = sum(!is.na(Area))/num)
# write.csv(aq.detect,"021617_in.cyano.dcm.diel.all.csv")
```

```{r combine}
measured.tqs <- aq.detect %>% filter(percent.dcm.blksub>0.25) %>%
     filter(percent.dcm.raw>11/num)

means.aq <- mydata_new %>%
     filter(!grepl("Blk",Replicate.Name)) %>%  
     filter(!grepl("Poo",Replicate.Name)) %>%
     select(Compound.Name, Area.BlkSub) %>%
     group_by(Compound.Name) %>%
     summarise(mean.dcm = mean(Area.BlkSub, na.rm=T))

measured.tqs <- full_join(measured.tqs, means.aq)
```

```{r aq.keepfewercompounds}
keep.me.aq <- measured.tqs %>%
     filter(!is.na(percent.dcm.blksub)) %>%
     select(Compound.Name)
keep.me.aq <- paste(c(keep.me.aq[[1]],IS.aq.list),collapse = "|")
aq.data.sub <- mydata_new %>%
     filter(grepl(keep.me.aq,Compound.Name))
```

##Back fill for compounds which we measure in at least 25% of samples
but don't measure for all samples
Output of this section is a dataframe aq.data.backfilled
that should have Updated values in the Area.BlkSub COlumn
and an added note if the data was 'backfilled'
```{r aq.backfill}
##select data that has no blank ratio
unfilled.BlkRatio <- aq.data.sub %>% 
     select(runDate, type, SampID, replicate, Compound.Name, 
            BlkRatio) %>% 
     spread(Compound.Name, BlkRatio) %>% 
     filter(type == "Smp")%>%
     gather(Compound.Name, BlkRatio, -SampID, -runDate,
            -replicate, -type) %>%
     filter(!grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(is.na(BlkRatio)) %>%
     mutate(newArea = NA, Replicate.Name = NA)

## get blank data
blk.data <- mydata %>%
     filter(grepl("Blk",Replicate.Name))

for (i in 1:nrow(unfilled.BlkRatio)){
     blkAreas <- blk.data %>% 
          filter(Compound.Name == unfilled.BlkRatio$Compound.Name[i],
                 grepl(unfilled.BlkRatio$runDate[i],Replicate.Name)) %>%
          select(Area)
     if(nrow(blkAreas)>0){
          blkArea <- max(blkAreas)
     } else {
          blkAreas <- blk.data %>% 
          filter(Compound.Name == unfilled.BlkRatio$Compound.Name[i]) %>%
          select(Area)
          blkArea <- max(blkAreas)
     }
     unfilled.BlkRatio$newArea[i] <- blkArea*0.5
     unfilled.BlkRatio$BlkRatio[i] <- 1
     unfilled.BlkRatio$Replicate.Name[i] <- 
          unfilled.BlkRatio[i,1:4] %>%
          paste(collapse="_")
}

aq.data.backfilled <- full_join(aq.data.sub, unfilled.BlkRatio) %>% 
     mutate(Area.BlkSub = ifelse(!is.na(newArea), 
                                 newArea, Area.BlkSub)) %>%
     mutate(Notes = ifelse(!is.na(newArea),
                           "BackFilled",Notes))
```

##remove 21-24 since they don't have the right extraction standards
Running this will depend on your samples!
```{r aq, eval = F}
new.data <- aq.data.backfilled %>% 
     mutate(Area.norm = ifelse(
          test = grepl(paste(IS.aq.list, collapse="|"), Compound.Name),
                               yes = Area,
                               no = Area.BlkSub))


new.data.2 <- new.data %>%
     filter(SampID!="21",SampID!="22",SampID!="23",SampID!="24")
```

##Calculate mean values for each IS
```{r ISmeans}
IS.aq.dat <- new.data.2[grepl(paste(IS.aq.list, collapse="|"),
                            new.data.2$Compound.Name),] %>% 
     select(Compound.Name, Area.norm, Replicate.Name) %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     filter(!grepl("Poo",Replicate.Name)) 
IS.aq.means <- IS.aq.dat %>% 
     group_by(Compound.Name) %>%
     summarize(ave = mean(Area.norm))
```

##Normalize to each internal Standard
Actually scaling/adjusting by dividing by IS peak area and multiplying by the average IS peak area.
Output is a dataframe 'area.norm' that has each peak adjusted
to each possible internal standards
```{r echo=FALSE}
wArea<- new.data.2 %>% filter(!is.na(Compound.Name)) %>%
     select(Replicate.Name,Compound.Name,Area.norm) %>%
     spread(key=Compound.Name, value=Area.norm) %>%
     as.data.frame

this.IS <- IS.aq.list[1]
area.norm <- wArea[,-1] %>% 
          sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                 names(wArea))]) %>%
          as_data_frame %>% mutate(Replicate.Name = wArea$Replicate.Name) %>%
          gather(Compound,Area_Norm, -Replicate.Name)
this.mean <- IS.aq.means %>% filter(Compound.Name==this.IS) %>% 
     select(ave) %>% as.numeric
area.norm <- area.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
key <- ncol(area.norm)
count <- length(which(!is.na(area.norm$Area_Norm)))/
               length(unique(area.norm$Replicate.Name))
names(area.norm)[key] <- paste(this.IS,"Norm.Area",sep=".")
# print(paste(1, this.IS, count, sep="-"))

for (i in 2:length(IS.aq.list)){
     this.IS <- IS.aq.list[i]
     if(length(wArea[,grep(this.IS, names(wArea))])!=0){
          this.norm <- wArea[,-1] %>% 
               sapply(FUN = function(x) x/wArea[,grep(this.IS,
                                                      names(wArea))]) %>%
               as_data_frame %>% 
               mutate(Replicate.Name = wArea$Replicate.Name) %>%
               gather(Compound,Area_Norm, -Replicate.Name)
          this.mean <- IS.aq.means %>% filter(Compound.Name==this.IS) %>% 
               select(ave) %>% as.numeric
          this.norm <- this.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
          key <- ncol(area.norm)
          area.norm[,key+1] <- this.norm$Area_Norm
          names(area.norm)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          # print(paste(i, this.IS, count, sep="-"))
     }
}
glimpse(area.norm)
```

Combine 'area.norm' dataframe with the full data in 'new.data.2'
```{r mergewData}
area.norm <- area.norm %>%
     mutate(Run.Cmpd = paste(area.norm$Replicate.Name,
                             area.norm$Compound)) %>%
     filter(!is.na(Replicate.Name)) %>%
     filter(!(Compound=="<NA>"))

new.data.2 <- new.data.2 %>%
     filter(!is.na(Replicate.Name)) %>%
     mutate(Run.Cmpd = ifelse(is.na(Run.Cmpd),
                    paste(new.data.2$Replicate.Name,
                          new.data.2$Compound.Name),
                    Run.Cmpd))

dat <- full_join(new.data.2, area.norm)
dat <- dat %>% select(-Compound) %>% filter(!is.na(type))
```

##Compare normalizations
get mean, sd, and rsd for each normalization of each compound for each sample type.
Output is a dataframe called 'rsd.stats'
```{r}
no.blank.dat <- dat %>% filter(type =="Smp" | type =="Poo") %>%
     filter(SampID!="B")

rsd.stats <- no.blank.dat %>% select(-Replicate.Name,
                                     -Retention.Time,
                                     -Notes, -Height, -IR, -S.N,
                                     -Run.Cmpd,
                                     -runDate, -replicate,
                                     -Area, -Area.BlkSub,
                                     -rawArea, 
                                     -newArea,
                                     -BlkRatio) %>%
     gather(Normer, Value, -Compound.Name, -type, - SampID) %>%
     group_by(Compound.Name, type, SampID, Normer) %>%
     summarize(m = mean(Value, na.rm=T), 
                sd = sd(Value, na.rm=T), rsd = sd/m)
glimpse(rsd.stats)
```

##Cleanup RSD data and add a minimizing column 
One based on the multiple injetions of the pooled sample "PooModel"
and one based on the injections of the biological replicates of samples "SmpModel"
```{r newModeMaker}
rsd.clean <- rsd.stats %>% filter(!is.na(m)) %>%
     filter(Normer!="Area.norm")
SmpModel <- rsd.clean %>% filter(type=="Smp") %>%
     select(-m, -sd) %>%
     group_by(Compound.Name, Normer) %>%
     summarize(Mean.rsd = mean(rsd, na.rm=T)) %>%
     summarize(SmpModelRSD = min(Mean.rsd),
               Smp.Picked.IS = unique(Normer)[which.min(Mean.rsd)])
PooModel <- rsd.clean %>% filter(type=="Poo") %>%
     select(-m, -sd) %>%
     group_by(Compound.Name, Normer) %>%
     summarize(Mean.rsd = mean(rsd, na.rm=T)) %>%
     filter(!is.na(Mean.rsd)) %>%
     summarize(PooModelRSD = min(Mean.rsd),
               Poo.Picked.IS = unique(Normer)[which.min(Mean.rsd)])
Models <- full_join(PooModel, SmpModel)
glimpse(Models)
```

##Merge Models and RSD data all together
```{r}
rsd.total <- full_join(rsd.stats, Models)
```

##Adjust the pooled model --> PooPlus model aka B-MIS model
so that it normalizes metabolites that we have internal standards
for to the appropriate internal standard. Otherwise keeps the pooled
model. 
Otherwise uses the PooModel for all compounds detected in the Pooled samples.
If a compound was detected in the sample but not the pooled then use the SmpModel.
Names this new complied model "PooPlus" (aka B-MIS)
Output of this section is 'rsd.total': A dataframe with mean, sd, and rsd for each compound, sample type, and normalization combination and a column indicating the poo, smp and pooPlus (B-MIS) internal standard selections 
```{r echo=FALSE}
rsd.total <- rsd.total %>%
     mutate(PooPlus.IS = Poo.Picked.IS) %>%
     mutate(PooPlusModelRSD = PooModelRSD) %>%
     ungroup() %>%
     mutate(Compound.Name = ifelse(Compound.Name=="Sulfolactate",
                                   yes = "Sulfolactic Acid",
                                   no = Compound.Name))


no.poo.rows <-which(is.na(rsd.total$PooModelRSD))
rsd.total$PooPlus.IS[no.poo.rows] <- rsd.total$Smp.Picked.IS[no.poo.rows]
rsd.total$PooPlusModelRSD[no.poo.rows] <- rsd.total$SmpModelRSD[no.poo.rows]

for (i in 1:nrow(rsd.total)){
     cmpd <- rsd.total$Compound.Name[i]
     if(length(grep(cmpd, IS.aq.list))>0){
          newIS <- paste0(IS.aq.list[grep(cmpd,
                                       IS.aq.list)],".Norm.Area")
          rsd.total$PooPlus.IS[i] <- newIS
     }
    
}
rsd.total <- rsd.total %>% 
     select(-PooModelRSD,-PooPlusModelRSD,-SmpModelRSD)
glimpse(rsd.total)
```

##Write out RSD data, if you want
```{r, eval=FALSE}
write.csv(rsd.total, "021617_Cyano_dcm_diel_all_normalization_rsd_dat.csv")
```

##Get Actual Normalized data not just the rsds
```{r, echo=FALSE}
models <- rsd.total %>% ungroup %>%
     select(Compound.Name,
            Poo.Picked.IS,Smp.Picked.IS,PooPlus.IS) %>%
     group_by(Compound.Name) %>%
     summarize(PooModel.IS = unique(Poo.Picked.IS),
               SmpModel.IS = unique(Smp.Picked.IS),
               PooPlusModel.IS = unique(PooPlus.IS))

dat <- dat %>% filter(!is.na(Replicate.Name)) %>%
     mutate(Compound.Name = ifelse(Compound.Name=="Sulfolactate",
                                   yes = "Sulfolactic Acid",
                                   no = Compound.Name))
dat.join <- full_join(dat, models)

PooModel <- c()
SmpModel <-c()
PooPlusModel <- c()
for (i in 1:nrow(dat.join)){
     if(!is.na(dat.join$PooModel.IS[i])){
          PooModel <- c(PooModel,dat.join[i,dat.join$PooModel.IS[i]])
     } else {
          PooModel <- c(PooModel,NA)
     }
     if(!is.na(dat.join$SmpModel.IS[i])){
          SmpModel <- c(SmpModel,dat.join[i,
                                            dat.join$SmpModel.IS[i]])
     } else {
          SmpModel <- c(SmpModel,NA)
     }
     if(!is.na(dat.join$PooPlusModel.IS[i])){
          PooPlusModel <- c(PooPlusModel,dat.join[i,
                                            dat.join$PooPlusModel.IS[i]])
     } else {
          PooPlusModel <- c(PooPlusModel,NA)
     }
}
dat.join$PooModel <- PooModel
dat.join$SmpModel <- SmpModel
dat.join$PooPlusModel <- PooPlusModel
dat.join <- dat.join %>% unnest(PooModel) %>% unnest(SmpModel) %>%
     unnest(PooPlusModel) %>%
     filter(!is.na(SampID))
glimpse(dat.join)
```
##Import cruise log
```{r cruiselog, echo=F}
cruiseLog <- read_csv("Clean Metabolomics Sample Log for SCOPE Cruise_DIEL.csv")
volData <- cruiseLog %>% select(`Sample ID`, Replicate, Vol, Date) %>%
     rename(replicate = Replicate, SampID = `Sample ID`) %>%
     mutate(SampID = as.character(SampID), 
            replicate = as.character(replicate))
```

##Merge data and vol data
and then correct area for volume filtered.
new.data is a dataframe with the column Area.norm that has
the updated blank subtracted values normalized to water volume.
```{r aq.areaCorrect}
final.data <- full_join(dat.join, volData)
final.data <- final.data %>% 
     mutate(Vol = ifelse(type=="Blk",3.5,
                         ifelse(type=="Poo", 3.5, Vol)))
final.data <- final.data %>%
     filter(!is.na(PooPlusModel)) %>% 
     mutate(PooPlusNormed =  PooPlusModel/Vol) %>%
     filter(!is.na(PooPlusModel))
```

##Write out that full normalized data
The output of the B-MIS normalized data is in the column named PooPlusModel.
```{r, eval=FALSE}
new.filename <- paste("021617_Normalized_data",filename,sep="_")
write.csv(final.data, new.filename)
```
