---
title: "normalize cyano aq diel"
author: "Angie Boysen"
date: "September 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(  
     message = FALSE,
     warning = FALSE,
     error = FALSE, 
     collapse = TRUE,
     comment = "#",
     fig.align = "center"
)

library(ggplot2)
library(gplots)
library(tidyr)
library(tibble)
require(graphics); require(grDevices)
library(Hmisc)
library(gtools)
library(cowplot)
require(RColorBrewer)
library(xlsx)
library(readr)
library(plotly)
library(stringr)
library(GGally)
library(ggbiplot)
library(stats)
library(dplyr)
```

##Exploring the cyano eddy samples

#Import data
#I'm going to start by processing all the data together
```{r echo=FALSE}
mydata.overloaded <- read_csv("QC_output_All_Diel_1_Cyano_Aq.csv",comment = "#")
aq.data <-mydata.overloaded[,2:9]
# dcm.data <- read_csv("QC_output2Diatom_cyanoDCM_160808.csv", comment = "#")
# dcm.data <- dcm.data[,2:9]
glimpse(aq.data)
```

##Read in Internal Standard list
```{r echo=FALSE}
IS.cat <- read_csv("../CYANO_MasterList_Summer2016.csv")  
IS.cat <- IS.cat %>% filter(Group == "Internal Std") %>% 
     select(`Compound Name`) %>% 
     unique()
IS.list <- IS.cat[[1]]
IS.aq.list <- IS.cat[[1]][c(1:4,6)]
IS.dcm.list <- IS.cat[[1]][c(3,6,8)]
```

##look at extraction replication
if things don't look good with some IS make sure to fix them
or at least not include them as possibilities for normalization
```{r aq}
IS.aq.dat <- aq.data[grepl(paste(IS.aq.list, collapse="|"),
                            aq.data$Compound.Name),]
IS.aq.dat <- IS.aq.dat %>% 
     select(Compound.Name, Area, Replicate.Name) %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     filter(!grepl("Poo",Replicate.Name)) 
ggplot(IS.aq.dat, aes(x=Replicate.Name, y=Area)) + 
     geom_bar(stat="identity") + 
     facet_wrap( ~Compound.Name, scales="free_y")+
     theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
           legend.position = "top")  + ggtitle("IS Raw Areas - in Aq")
```

##Edit data 
so d-Phenylalanine, B2-IS are halved in 6R1
And so 
```{r correctISErrors}
aq.data.fixed <- aq.data
errors <- aq.data %>%
     filter(Replicate.Name=="160701_Smp_6R1") %>% 
     filter(Compound.Name=="D-Phenylalanine"|
                 Compound.Name=="Vitamin B2_IS") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name==
                         "160701_Smp_6R1") &
                  (aq.data.fixed$Compound.Name==
                        "D-Phenylalanine"|
                 aq.data.fixed$Compound.Name==
                      "Vitamin B2_IS")
             ] <- errors$Area

errors <- aq.data %>%
     filter(Replicate.Name=="160701_Smp_5R1") %>% 
     filter(Compound.Name=="D-Tryptophan"|
                 Compound.Name=="d-IAA"|
                 Compound.Name=="d4-Tryptamine") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name==
                         "160701_Smp_5R1") &
                  (aq.data.fixed$Compound.Name==
                        "D-Tryptophan"|
                 aq.data.fixed$Compound.Name==
                      "d-IAA"|
                      aq.data.fixed$Compound.Name==
                      "d4-Tryptamine")
             ] <- errors$Area
```

##Pull out compounds that were detected
```{r detected?, echo=F}
aq.detect <- aq.data.fixed %>%      
     filter(!grepl("Blk",Replicate.Name)) %>%
     select(Replicate.Name, Compound.Name, Area) %>%
     group_by(Compound.Name) %>%
     summarise(percent.aq = sum(!is.na(Area))/n())
# write.csv(aq.detect,"in.cyano.aq.diel.csv")
```

##Rename Runs
Name structure must be:
Date_type_ID_replicate
```{r echo=FALSE}
aq.data.fixed <- aq.data.fixed %>% 
     mutate(Replicate.Name = Replicate.Name %>%
                 str_replace("R1",
                             "_1") %>%
                 str_replace("R2",
                             "_2") %>%
                 str_replace("R3",
                             "_3") %>%
                 str_replace("160701_Blk_6",
                             "160701_Blk_Blk_6") %>%
                 str_replace("160701_Blk_15",
                             "160701_Blk_Blk_15") %>%
                 str_replace("160708_Blk_16",
                             "160708_Blk_Blk_16") %>%
                 str_replace("160708_Blk_5",
                             "160708_Blk_Blk_5") %>%
                 str_replace("160714_Blk_12",
                             "160714_Blk_Blk_12") %>%
                 str_replace("160714_Blk_4",
                             "160714_Blk_Blk_4") %>%
                 str_replace("160714_Blk_7-11_FilterBlk",
                             "160714_Blk_FilterBlk_7-11"))
```

##Calculate mean values for each IS
```{r ISmeans}
IS.aq.means <- IS.aq.dat %>% 
     filter(!grepl("Blk", Replicate.Name)) %>%
     group_by(Compound.Name) %>%
     summarize(ave = mean(Area))
```

##Normalize to each internal Standard for which there aren't NAs
Actually scaling/adjusting by dividing by IS peak area and multiplying by the average IS peak area
```{r aqNormalization, echo=FALSE}
##Clean up data - remove rows that only have Area in the blank samples
wArea.aq <- aq.data.fixed %>%
     select(Replicate.Name,Compound.Name,Area) %>%
     spread(key=Compound.Name, value=Area) %>% 
     filter(!grepl("Blk",Replicate.Name)) %>%
     Filter(function(x) !all(is.na(x)), .) %>%
     as.data.frame

## Normalize to ISs
this.IS <- IS.aq.means[[1,1]]
area.norm.aq <- wArea.aq[,-1] %>% 
     sapply(FUN = function(x) x/wArea.aq[,grep(this.IS,
                                               names(wArea.aq))]) %>%
     as_data_frame %>% 
     mutate(Replicate.Name = wArea.aq$Replicate.Name) %>%
     gather(Compound,Area_Norm, -Replicate.Name)
this.mean <- IS.aq.means %>% filter(Compound.Name==this.IS) %>% 
     select(ave) %>% as.numeric
area.norm.aq <- area.norm.aq %>% mutate(Area_Norm = Area_Norm*this.mean)
key <- ncol(area.norm.aq)
count <- length(which(!is.na(area.norm.aq$Area_Norm)))/
               length(unique(area.norm.aq$Replicate.Name))
names(area.norm.aq)[key] <- paste(this.IS,"Norm.Area",sep=".")
print(paste(1, this.IS, count, sep="-"))

for (i in 2:length(IS.aq.list)){
     if(!is.na(IS.aq.means[[i,2]])){
     this.IS <- IS.aq.means[[i,1]]
     if(length(wArea.aq[,grep(this.IS, names(wArea.aq))])!=0){
          this.norm <- wArea.aq[,-1] %>% 
               sapply(FUN = function(x) x/wArea.aq[,grep(this.IS,
                                                      names(wArea.aq))]) %>%
               as_data_frame %>% 
               mutate(Replicate.Name = wArea.aq$Replicate.Name) %>%
               gather(Compound,Area_Norm, -Replicate.Name)
          this.mean <- IS.aq.means %>% filter(Compound.Name==this.IS) %>% 
               select(ave) %>% as.numeric
          this.norm <- this.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
          key <- ncol(area.norm.aq)
          area.norm.aq[,key+1] <- this.norm$Area_Norm
          names(area.norm.aq)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          print(paste(i, this.IS, count, sep="-"))
     }
     }
}
```

#Break Up the Names
Name structure must be:
Date_type_ID_replicate
```{r echo=FALSE}
mydata_new <- area.norm.aq %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type",
                                        "SampID",
                                        "replicate"),"_") %>%
     mutate(Run.Cmpd =
                 paste(area.norm.aq$Replicate.Name,
                       area.norm.aq$Compound))
aq.data.fixed <- aq.data.fixed %>% 
     mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))

aq.dat <- full_join(aq.data.fixed, mydata_new)
aq.dat <- aq.dat %>% select(-Compound)
glimpse(aq.dat)
```

##Compare normalizations
get rsd for each normalization of each compound for each sample
```{r aq, echo=FALSE}
no.blank.dat.aq <- aq.dat %>% filter(type =="Smp" | type =="Poo") 
rsd.stats.aq <- no.blank.dat.aq %>% select(-Replicate.Name,
                                           -Retention.Time,
                                           -Notes, -Height, -IR, -S.N,
                                           -Run.Cmpd,
                                           -runDate, -replicate) %>%
     gather(Normer, Value, -Compound.Name, -type, - SampID) %>%
     group_by(Compound.Name, type, SampID, Normer) %>%
     summarize(m = mean(Value, na.rm=T), 
                sd = sd(Value, na.rm=T), rsd = sd/m) %>%
     filter(!is.na(m))
```

##Cleanup RSD data and add a minimizing column 
#==> dynamic model of normalization
```{r newModeMaker}
ModelMaker <- function(rsd.stats){
     rsd.clean <- rsd.stats %>% filter(!is.na(rsd)) %>%
          filter(Normer!="Area")
     SmpModel <- rsd.clean %>% filter(type=="Smp") %>%
          select(-m, -sd) %>%
          group_by(Compound.Name, Normer) %>%
          summarize(Mean = mean(rsd, na.rm=T)) %>%
          summarize(SmpModelRSD = min(Mean),
                    Smp.Picked.IS = unique(Normer)[which.min(Mean)])
     PooModel <- rsd.clean %>% filter(type=="Poo") %>%
          select(-m, -sd) %>%
          group_by(Compound.Name, Normer) %>%
          summarize(Mean = mean(rsd, na.rm=T)) %>%
          summarize(PooModelRSD = min(Mean),
                    Poo.Picked.IS = unique(Normer)[which.min(Mean)])
     Models <- full_join(PooModel, SmpModel)
     Models
}
```

```{r}
models.aq <- ModelMaker(rsd.stats.aq)
```

##Merge Models and RSD data all together
```{r aq, echo=FALSE}
rsd.total.aq <- full_join(rsd.stats.aq,models.aq)
```

##Adjust the pooled model 
so that it normalizes metabolites that we have internal standards for to 
the appropriate internal standard. Otherwise keeps the pooled model. Names this new model "PooPlus"
```{r aq}
rsd.total.aq <- rsd.total.aq %>%
     mutate(PooPlus.IS = Poo.Picked.IS) %>%
     mutate(PooPlusModelRSD = PooModelRSD)
used.IS <- IS.aq.means %>% filter(!is.na(ave)) %>% select(Compound.Name)
used.IS <- used.IS[[1]]     

for (i in 1:nrow(rsd.total.aq)){
     cmpd <- rsd.total.aq$Compound.Name[i]
     if(length(grep(cmpd, used.IS))>0){
          newIS <- paste0(IS.list[grep(cmpd, IS.list)],".Norm.Area")
          rsd.total.aq$PooPlus.IS[i] <- newIS
          rsd.total.aq$PooPlusModelRSD[i] <- filter(rsd.total.aq, 
                                                 Compound.Name == cmpd, 
                                                 Normer == newIS, 
                                                 type == "Poo")$rsd
     }
     ## if the poo didn't pick a normalizing IS, use the samp picked one
     if(is.na(rsd.total.aq$Poo.Picked.IS[i]) &
        !is.na(rsd.total.aq$Smp.Picked.IS[i])){
          rsd.total.aq$PooPlus.IS[i] <- rsd.total.aq$Smp.Picked.IS[i]
          rsd.total.aq$PooPlusModelRSD[i] <-
               rsd.total.aq$SmpModelRSD[i]
     }
}
```

##Write rsd data
```{r eval=FALSE}
write.csv(rsd.total.aq, "Cyano_Aq_diel_1_normalization_rsd_dat_only2IS.csv")
# write.csv(rsd.total.dcm, "Cyano_DCM_diatom_normalization_rsd_dat.csv")
```

##Get Actual Normalized data not just the rsds
```{r aq, echo=FALSE}
models.aq <- rsd.total.aq %>% ungroup %>%
     select(Compound.Name,
            Poo.Picked.IS,Smp.Picked.IS,PooPlus.IS) %>%
     group_by(Compound.Name) %>%
     summarize(PooModel.IS = unique(Poo.Picked.IS),
               SmpModel.IS = unique(Smp.Picked.IS),
               PooPlusModel.IS = unique(PooPlus.IS))
aq.dat <- aq.dat %>% filter(!is.na(Replicate.Name))
dat.join.aq <- full_join(aq.dat, models.aq)

PooModel <- c()
SmpModel <-c()
PooPlusModel <- c()
for (i in 1:nrow(dat.join.aq)){
     if(!is.na(dat.join.aq$PooModel.IS[i])){
          PooModel <- c(PooModel,dat.join.aq[i,dat.join.aq$PooModel.IS[i]])
     } else {
          PooModel <- c(PooModel,NA)
     }
     if(!is.na(dat.join.aq$SmpModel.IS[i])){
          SmpModel <- c(SmpModel,dat.join.aq[i,
                                            dat.join.aq$SmpModel.IS[i]])
     } else {
          SmpModel <- c(SmpModel,NA)
     }
     if(!is.na(dat.join.aq$PooPlusModel.IS[i])){
          PooPlusModel <- c(PooPlusModel,dat.join.aq[i,
                                            dat.join.aq$PooPlusModel.IS[i]])
     } else {
          PooPlusModel <- c(PooPlusModel,NA)
     }
}
dat.join.aq$PooModel <- PooModel
dat.join.aq$SmpModel <- SmpModel
dat.join.aq$PooPlusModel <- PooPlusModel

dat.join.aq <- dat.join.aq %>% unnest(PooModel) %>% unnest(SmpModel) %>%
     unnest(PooPlusModel)
```

dcm 
```{r dcm, echo=FALSE}
models.dcm <- rsd.total.dcm %>% ungroup %>%
     select(Compound.Name,
            Poo.Picked.IS,Smp.Picked.IS,PooPlus.IS) %>%
     group_by(Compound.Name) %>%
     summarize(PooModel.IS = unique(Poo.Picked.IS),
               SmpModel.IS = unique(Smp.Picked.IS),
               PooPlusModel.IS = unique(PooPlus.IS))
DCM.dat <- DCM.dat %>% filter(!is.na(Replicate.Name))
dat.join.dcm <- full_join(DCM.dat, models.dcm)

PooModel <- c()
SmpModel <-c()
PooPlusModel <- c()
for (i in 1:nrow(dat.join.dcm)){
     if(!is.na(dat.join.dcm$PooModel.IS[i])){
          PooModel <- c(PooModel,dat.join.dcm[i,dat.join.dcm$PooModel.IS[i]])
     } else {
          PooModel <- c(PooModel,NA)
     }
     if(!is.na(dat.join.dcm$SmpModel.IS[i])){
          SmpModel <- c(SmpModel,dat.join.dcm[i,
                                            dat.join.dcm$SmpModel.IS[i]])
     } else {
          SmpModel <- c(SmpModel,NA)
     }
     if(!is.na(dat.join.dcm$PooPlusModel.IS[i])){
          PooPlusModel <- c(PooPlusModel,dat.join.dcm[i,
                                            dat.join.dcm$PooPlusModel.IS[i]])
     } else {
          PooPlusModel <- c(PooPlusModel,NA)
     }
}
dat.join.dcm$PooModel <- PooModel
dat.join.dcm$SmpModel <- SmpModel
dat.join.dcm$PooPlusModel <- PooPlusModel

dat.join.dcm <- dat.join.dcm %>% unnest(PooModel) %>% unnest(SmpModel) %>%
     unnest(PooPlusModel)
```

##Write out that full normalized data
```{r}
write.csv(dat.join.aq, "CyanoAq_Diatom_Samples_Normalized_allTheWays.new0905.csv")
write.csv(dat.join.dcm, "CyanoDCM_Diatom_Samples_Normalized_allTheWays.new0905.csv")
```

##scatter plots
```{r echo=F}
samp.dat <- dat.join.aq %>%
     filter(Compound.Name!="D-Tryptophan"&
                 Compound.Name!="D-Phenylalanine"&
                 Compound.Name!="d-IAA"&
                 Compound.Name!="d4-Tryptamine"&
                 Compound.Name!="Vitamin B2_IS"&
                 Compound.Name!="D8-Arachidonic Acid") %>%
     select(Compound.Name,Area,SampID,
            replicate,`D-Phenylalanine.Norm.Area`,
            `Vitamin B2_IS.Norm.Area`, PooModel, SmpModel) %>% 
     gather(Norm,RelArea,-Compound.Name,-SampID,-replicate) %>%
     filter(replicate==34 | replicate==12 | replicate==56)

samp.means <- samp.dat %>%
     group_by(Compound.Name, SampID, Norm) %>%
     summarize(Mean = mean(RelArea)) %>%
     spread(SampID,Mean)

samp.sds <- samp.dat %>% 
     group_by(Compound.Name, SampID, Norm) %>%
     summarize(sd = sd(RelArea)) %>%
     spread(SampID,sd) %>%
     rename(B.sd = B, Np.sd =Np, Pt.sd=Pt,Tp.sd=Tp)
samp.pval <- samp.dat %>% 
     spread(SampID,RelArea) %>%
     group_by(Compound.Name,Norm) %>%
     summarize(Np.Pt.p =
                    ifelse(!is.na(mean(Np))&!is.na(mean(Pt)),
                           t.test(Np,Pt,alt="two.sided")$p.value, 
                           NA),
               Np.Tp.p =
                    ifelse(!is.na(mean(Np))&!is.na(mean(Tp)),
                           t.test(Np,Tp,alt="two.sided")$p.value, 
                           NA),
               Pt.Tp.p =
                    ifelse(!is.na(mean(Pt))&!is.na(mean(Tp)),
                           t.test(Pt,Tp,alt="two.sided")$p.value, 
                           NA)) %>%
     mutate(Pt.Tp.p = Pt.Tp.p<0.05,
            Np.Tp.p = Np.Tp.p<0.05,
            Np.Pt.p = Np.Pt.p<0.05)

samp.summary <- full_join(samp.sds, samp.means)
samp.summary <- full_join(samp.summary, samp.pval)

Pt.Tp <- ggplot(samp.summary, aes(x=Pt, y=Tp, text=Compound.Name)) + geom_point(aes(color=Pt.Tp.p),alpha=0.9,size=2) + scale_color_hue(l=45)+ facet_wrap(~Norm,scales="free") +  geom_errorbar(aes(ymin = Tp-Tp.sd, ymax = Tp+Tp.sd)) + geom_errorbarh(aes(xmin = Pt-Pt.sd, xmax = Pt+Pt.sd)) + geom_abline(slope = 1, intercept = 0) + scale_x_log10() + scale_y_log10() +ggtitle("Pt vs Tp")
ggplotly(Pt.Tp)

Tp.Np <- ggplot(samp.summary, aes(x=Np, y=Tp, text=Compound.Name)) + geom_point(aes(color=Np.Tp.p),alpha=0.9,size=2) + scale_color_hue(l=45)+ facet_wrap(~Norm,scales="free") +  geom_errorbar(aes(ymin = Tp-Tp.sd, ymax = Tp+Tp.sd)) + geom_errorbarh(aes(xmin = Np-Np.sd, xmax = Np+Np.sd)) + geom_abline(slope = 1, intercept = 0) + scale_x_log10() + scale_y_log10() +ggtitle("Np vs Tp")
ggplotly(Tp.Np)
```

stat tests
```{r stats, echo=F}
##Analysis of variance (aov)
dmb.poo <- dmbz %>% filter(norm=="raw.TIC")
dmb.poo.aov <- aov(Relative.Area~description, data = dmb.poo)
summary(dmb.poo.aov)
TukeyHSD(dmb.poo.aov)

syr.wide <- Syringaldehyde %>% spread(description,Relative.Area) %>% filter(norm=="poo.norm")
wilcox.test(syr.wide$Eddy10,syr.wide$Eddy12, alt="less")
test<-t.test(syr.wide$Eddy10,syr.wide$Eddy12,alt="less")

```
