---
title: "normalize cyano aq diel"
author: "Angie Boysen"
date: "September 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(  
     message = FALSE,
     warning = FALSE,
     error = FALSE, 
     collapse = TRUE,
     comment = "#",
     fig.align = "center"
)

library(ggplot2)
library(gplots)
library(tidyr)
library(tibble)
require(graphics); require(grDevices)
library(Hmisc)
library(gtools)
library(cowplot)
require(RColorBrewer)
library(xlsx)
library(readr)
library(plotly)
library(stringr)
library(GGally)
library(ggbiplot)
library(stats)
library(dplyr)
```

##Exploring the cyano eddy samples

#Import data
#I'm going to start by processing all the data together
```{r echo=FALSE}
mydata.overloaded <- read_csv("QC_output_wBlkSub_All_Diel_1_Cyano_Aq.csv",comment = "#")
aq.data <-mydata.overloaded[,2:11]
dcm.data <- read_csv("QC_output_wBlkSubDiel_1_Cyano_DCM_all.csv", comment = "#")
dcm.data <- dcm.data[,2:11]
glimpse(aq.data)
```

##Read in Internal Standard list
```{r echo=FALSE}
IS.cat <- read_csv("../CYANO_MasterList_Summer2016.csv")  
IS.cat <- IS.cat %>% filter(Group == "Internal Std") %>% 
     select(`Compound Name`) %>% 
     unique()
IS.list <- IS.cat[[1]]
IS.aq.list <- IS.cat[[1]][c(1:4,6)]
IS.dcm.list <- IS.cat[[1]][c(3,6,8)]
```


##look at extraction replication
if things don't look good with some IS make sure to fix them
or at least not include them as possibilities for normalization
```{r aq}
IS.aq.dat <- aq.data[grepl(paste(IS.aq.list, collapse="|"),
                            aq.data$Compound.Name),]
IS.aq.dat <- IS.aq.dat %>% 
     select(Compound.Name, Area, Replicate.Name) %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     filter(!grepl("Poo",Replicate.Name)) 
ggplot(IS.aq.dat, aes(x=Replicate.Name, y=Area)) + 
     geom_bar(stat="identity") + 
     facet_wrap( ~Compound.Name, scales="free_y")+
     theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
           legend.position = "top")  + ggtitle("IS Raw Areas - in Aq")
```


```{r dcm}
IS.dcm.dat <- dcm.data[grepl(paste(IS.dcm.list, collapse="|"),
                             dcm.data$Compound.Name),]
IS.dcm.dat <- IS.dcm.dat %>% select(Compound.Name, Area, Replicate.Name)
ggplot(IS.dcm.dat, aes(x=Replicate.Name, y=Area)) + 
     geom_bar(stat="identity") + 
     facet_wrap( ~Compound.Name, scales="free_y")+
     theme(axis.text.x = element_text(angle = 90, hjust = 1,vjust = 0.5),
           legend.position = "top")  + ggtitle("IS Raw Areas - in DCM")
```


##Edit data 
so d-Phenylalanine, B2-IS are halved in 6R1
And so 5R1 has half the D-Tryptophan, d4-Tryptamine, d-IAA
```{r correctISErrors}
aq.data.fixed <- aq.data
errors <- aq.data %>%
     filter(Replicate.Name=="160701_Smp_6R1") %>% 
     filter(Compound.Name=="D-Phenylalanine"|
                 Compound.Name=="Vitamin B2_IS") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name==
                         "160701_Smp_6R1") &
                  (aq.data.fixed$Compound.Name==
                        "D-Phenylalanine"|
                 aq.data.fixed$Compound.Name==
                      "Vitamin B2_IS")
             ] <- errors$Area

errors <- aq.data %>%
     filter(Replicate.Name=="160701_Smp_5R1") %>% 
     filter(Compound.Name=="D-Tryptophan"|
                 Compound.Name=="d-IAA"|
                 Compound.Name=="d4-Tryptamine") %>%
     mutate(Area = Area/2)
aq.data.fixed$Area[(aq.data.fixed$Replicate.Name==
                         "160701_Smp_5R1") &
                  (aq.data.fixed$Compound.Name==
                        "D-Tryptophan"|
                 aq.data.fixed$Compound.Name==
                      "d-IAA"|
                      aq.data.fixed$Compound.Name==
                      "d4-Tryptamine")
             ] <- errors$Area
```

##Rename Runs
Name structure must be:
Date_type_ID_replicate
```{r aq echo=FALSE}
aq.data.fixed <- aq.data %>% 
     mutate(Replicate.Name = Replicate.Name %>%
                 str_replace("R1",
                             "_1") %>%
                 str_replace("R2",
                             "_2") %>%
                 str_replace("R3",
                             "_3") %>%
                 str_replace("160701_Blk_6",
                             "160701_Blk_Blk_6") %>%
                 str_replace("160701_Blk_15",
                             "160701_Blk_Blk_15") %>%
                 str_replace("160708_Blk_16",
                             "160708_Blk_Blk_16") %>%
                 str_replace("160708_Blk_5",
                             "160708_Blk_Blk_5") %>%
                 str_replace("160714_Blk_12",
                             "160714_Blk_Blk_12") %>%
                 str_replace("160714_Blk_4",
                             "160714_Blk_Blk_4") %>%
                 str_replace("160714_Blk_7-11_FilterBlk",
                             "160714_Blk_FilterBlk_7-11") %>%
                 str_replace("160701_Poo_AllCyanoAqExtracts",
                             "160701_Poo_AllCyanoAqExtracts1")%>%
                 str_replace("160708_Poo_AllCyanoAqExtracts",
                             "160708_Poo_AllCyanoAqExtracts2")%>%
                 str_replace("160714_Poo_AllCyanoAqExtracts",
                             "160714_Poo_AllCyanoAqExtracts3"))
```

```{r dcm, echo=F}
dcm.data.fixed <- dcm.data %>%
     mutate(Replicate.Name = Replicate.Name %>%
                 str_replace("R1",
                             "_1") %>%
                 str_replace("R2",
                             "_2") %>%
                 str_replace("R3",
                             "_3") %>%
                 str_replace("_DCM",
                             "" ) %>%
                 str_replace("AllCyanoDCM-Extracts",
                             "AllCyanoDCMExtracts3") %>%
                 str_replace("160701_Poo_AllCyanoDCMExtracts",
                             "160701_Poo_AllCyanoDCMExtracts1") %>%
                 str_replace("160708_Poo_AllCyanoDCMExtracts",
                             "160708_Poo_AllCyanoDCMExtracts2") %>%
                 str_replace("160714_Blk_6",
                             "160714_Blk_Blk_6") %>%
                 str_replace("160714_Blk_13",
                             "160714_Blk_Blk_13")%>%
                 str_replace("160708_Blk_16",
                             "160708_Blk_Blk_16") %>%
                 str_replace("160708_Blk_4",
                             "160708_Blk_Blk_4") %>%
                 str_replace("160701_Blk_16",
                             "160701_Blk_Blk_16")%>%
                 str_replace("160701_Blk_5",
                             "160701_Blk_Blk_5") %>%
                 
                 str_replace("160714_Blk_7-11_FilterBlk",
                             "160714_Blk_FilterBlk_7-11"))
                 
```


##Break up the names
Name structure must be:
Date_type_ID_replicate
```{r aq}
aq.data.fixed <- aq.data.fixed %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type",
                                        "SampID",
                                        "replicate"),"_") %>%
     mutate(Run.Cmpd =
                 paste(aq.data.fixed$Replicate.Name,
                       aq.data.fixed$Compound.Name),
            Replicate.Name = aq.data.fixed$Replicate.Name)
```

```{r dcm}
dcm.data.fixed <- dcm.data.fixed %>% separate(Replicate.Name, 
                                      c("runDate",
                                        "type",
                                        "SampID",
                                        "replicate"),"_") %>%
     mutate(Run.Cmpd =
                 paste(dcm.data.fixed$Replicate.Name,
                       dcm.data.fixed$Compound.Name),
            Replicate.Name = dcm.data.fixed$Replicate.Name)
```

#remove negatives
```{r aq.noNegs}
aq.data.fixed <- aq.data.fixed %>% 
     filter(Area.BlkSub>100 | 
                 is.na(Area.BlkSub) |
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(BlkRatio<0.5 |
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("S/N",Notes)|
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>% 
     filter(!grepl("too small",Notes)|
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("No second trace",Notes) |
                 grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("Bad RT",Notes))
```

```{r dcm.noNegs}
dcm.data.noNegs <- dcm.data.fixed %>% 
     filter(Area.BlkSub>100 | 
                 is.na(Area.BlkSub) |
                 grepl(paste(IS.dcm.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(BlkRatio<0.5 |
                 grepl(paste(IS.dcm.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("S/N",Notes)|
                 grepl(paste(IS.dcm.list, collapse="|"), 
                       Compound.Name)) %>% 
     filter(!grepl("too small",Notes)|
                 grepl(paste(IS.dcm.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("No second trace",Notes) |
                 grepl(paste(IS.dcm.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(!grepl("Bad RT",Notes)) %>%
     filter(Compound.Name!="Pyridoxamine")
```

##Pull out compounds that were detected
```{r aq.etected?, echo=F}
aq.detect <- aq.data.fixed %>%      
     filter(!grepl("Blk",Replicate.Name)) %>%
     select(Replicate.Name, Compound.Name, Area.BlkSub) %>%
     group_by(Compound.Name) %>%
     summarise(percent.aq = sum(!is.na(Area.BlkSub))/81)
# write.csv(aq.detect,"in.cyano.aq.diel.csv")
```

```{r dcm.detected?, echo=F}
DCM.detect <- dcm.data.noNegs %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     select(Replicate.Name, Compound.Name, Area.BlkSub) %>%
     group_by(Compound.Name) %>%
     summarize(percent.DCM = sum(!is.na(Area.BlkSub))/81)
# write.csv(DCM.detect,"in.cyano.DCM.diel.blksubdata.csv")
```

```{r combine}
aq.detect2 <- aq.detect %>% filter(percent.aq>0.2)
DCM.detect2 <- DCM.detect %>% filter(percent.DCM>0.2)
measured.tqs <- full_join(aq.detect2,DCM.detect2)

means.aq <- aq.data.fixed %>%
     filter(!grepl("Blk",Replicate.Name)) %>%  
     filter(!grepl("Poo",Replicate.Name)) %>%
     select(Compound.Name, Area.BlkSub) %>%
     group_by(Compound.Name) %>%
     summarise(mean.aq = mean(Area.BlkSub, na.rm=T))

means.org <- dcm.data.noNegs %>%
     filter(!grepl("Blk",Replicate.Name)) %>%  
     filter(!grepl("Poo",Replicate.Name)) %>%
     select(Compound.Name, Area.BlkSub) %>%
     group_by(Compound.Name) %>%
     summarise(mean.dcm = mean(Area.BlkSub, na.rm=T))

measured.tqs <- full_join(measured.tqs, means.aq)
measured.tqs <- full_join(measured.tqs, means.org) %>%
     mutate(a.over.o  = mean.aq/mean.dcm) %>%
     filter(!is.na(percent.aq) | !is.na(percent.DCM))

measured.tqs1 <- measured.tqs %>%
     mutate(pick = ifelse(is.na(percent.DCM),"a",
                          ifelse(is.na(percent.aq),"o",
                          ifelse(percent.aq ==0 & percent.DCM>0, "o",
                          ifelse(percent.aq>0 & percent.DCM==0,"a",
                          ifelse(percent.aq>percent.DCM & a.over.o>1,"a",
                          ifelse(percent.aq<percent.DCM & a.over.o<1,"o",
                          ifelse(a.over.o > 1,"a",
                          ifelse(abs(a.over.o) < 1, "o","tbd")))))))))

# write.csv(measured.tqs1, "in.cyano.diff.phases.csv")
```

##select only compounds that are tagged for that phase
or: If the compounds were found in less than 25% of samples, discard the data
```{r aq.keepfewercompounds}
# getRidOfMe <- aq.detect %>% filter(percent.aq<0.25) %>%
     # select(Compound.Name)
# getRidOfMe <- paste(getRidOfMe[[1]], collapse="|")
# aq.data.sub <- aq.data.fixed %>%
     # filter(!grepl(getRidOfMe, Compound.Name))

keep.me.aq <- measured.tqs1 %>% filter(pick=="a") %>%
     select(Compound.Name)
keep.me.aq <- paste(c(keep.me.aq[[1]],IS.aq.list),collapse = "|")
aq.data.sub <- aq.data.fixed %>%
     filter(grepl(keep.me.aq,Compound.Name))
```

```{r dcm.keepfewrcompounds}
# getRidOfMe.dcm <- DCM.detect %>% filter(percent.DCM<0.25) %>%
#      select(Compound.Name)
# getRidOfMe.dcm <- paste(getRidOfMe.dcm[[1]], collapse="|")
# dcm.data.sub <- dcm.data.noNegs %>%
#      filter(!grepl(getRidOfMe.dcm, Compound.Name)|grepl(paste(IS.dcm.list, collapse="|"), Compound.Name))

keep.me.dcm <- measured.tqs1 %>% filter(pick=="o") %>%
     select(Compound.Name)
keep.me.dcm <- paste(c(keep.me.dcm[[1]],IS.dcm.list),collapse = "|")
dcm.data.sub <- dcm.data.fixed %>%
     filter(grepl(keep.me.dcm,Compound.Name))
```

##Back fill for compounds which we measure in at least 25% of samples
but don't measure for all samples
```{r aq.backfill}
unfilled.BlkRatio <- aq.data.sub %>% 
     select(runDate, type, SampID, replicate, Compound.Name, 
            BlkRatio) %>% 
     spread(Compound.Name, BlkRatio) %>% 
     filter(type == "Smp")%>%
     gather(Compound.Name, BlkRatio, -SampID, -runDate, -replicate,
            -type) %>%
     filter(!grepl(paste(IS.aq.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(is.na(BlkRatio)) %>%
     mutate(newArea = NA, Replicate.Name = NA)
blk.data <- aq.data %>%
     filter(grepl("Blk",Replicate.Name))

for (i in 1:nrow(unfilled.BlkRatio)){
     blkAreas <- blk.data %>% 
          filter(Compound.Name ==unfilled.BlkRatio$Compound.Name[i],
                 grepl(unfilled.BlkRatio$runDate[i],Replicate.Name)) %>%
          select(Area)
     blkArea <- max(blkAreas)
     unfilled.BlkRatio$newArea[i] <- blkArea*0.5
     unfilled.BlkRatio$BlkRatio[i] <- 1
     unfilled.BlkRatio$Replicate.Name[i] <- 
          unfilled.BlkRatio[i,1:4] %>%
          paste(collapse="_")
}

aq.data.backfilled <- full_join(aq.data.sub, unfilled.BlkRatio) %>% 
     mutate(Area.BlkSub = ifelse(!is.na(newArea), 
                                 newArea, Area.BlkSub)) %>%
     mutate(Notes = ifelse(!is.na(newArea),
                           "BackFilled",Notes))
```

```{r dcm.backfill}
dcm.unfilled.BlkRatio <- dcm.data.sub %>% 
     select(runDate, type, SampID, replicate, Compound.Name, 
            BlkRatio) %>% 
     spread(Compound.Name, BlkRatio) %>% 
     filter(type == "Smp")%>%
     gather(Compound.Name, BlkRatio, -SampID, -runDate, -replicate,
            -type) %>%
     filter(!grepl(paste(IS.dcm.list, collapse="|"), 
                       Compound.Name)) %>%
     filter(is.na(BlkRatio)) %>%
     mutate(newArea = NA, Replicate.Name = NA)
dcm.blk.data <- dcm.data %>%
     filter(grepl("Blk",Replicate.Name))

for (i in 1:nrow(dcm.unfilled.BlkRatio)){
     blkAreas <- dcm.blk.data %>% 
          filter(Compound.Name ==dcm.unfilled.BlkRatio$Compound.Name[i],
                 grepl(dcm.unfilled.BlkRatio$runDate[i],Replicate.Name)) %>%
          select(Area)
     blkArea <- max(blkAreas)
     dcm.unfilled.BlkRatio$newArea[i] <- blkArea*0.5
     # dcm.unfilled.BlkRatio$BlkRatio[i] <- 1
     dcm.unfilled.BlkRatio$Replicate.Name[i] <- 
          dcm.unfilled.BlkRatio[i,1:4] %>%
          paste(collapse="_")
}

dcm.data.backfilled <- full_join(dcm.data.sub, dcm.unfilled.BlkRatio) %>% 
     mutate(Area.BlkSub = ifelse(!is.na(newArea), 
                                 newArea, Area.BlkSub)) %>%
     mutate(Notes = ifelse(!is.na(newArea),
                           "BackFilled",Notes)) %>%
     mutate(BlkRatio = ifelse(!is.na(newArea),1,BlkRatio))
```

##Import cruise log
```{r cruiselog, echo=F}
cruiseLog <- read_csv("Clean Metabolomics Sample Log for SCOPE Cruise_DIEL.csv")
volData <- cruiseLog %>% select(`Sample ID`, Replicate, Vol, Date) %>%
     filter(`Sample ID` <= 24) %>%
     rename(replicate = Replicate, SampID = `Sample ID`) %>%
     mutate(SampID = as.character(SampID), 
            replicate = as.character(replicate))
```

##Merge data and vol data
#and then correct area for volume filtered
```{r aq.areaCorrect}
new.data <- full_join(aq.data.backfilled, volData)
new.data <- new.data %>% 
     mutate(Vol = ifelse(type=="Blk",3.5,
                         ifelse(type=="Poo", 3.5, Vol)))
new.data <- new.data %>% 
     mutate(Area.norm = ifelse(
          test = grepl(paste(IS.aq.list, collapse="|"), Compound.Name),
                               yes = Area/Vol,
                               no = Area.BlkSub/Vol))
```

```{r dcm.areaCorrect}
dcm.new.data <- full_join(dcm.data.backfilled, volData)
dcm.new.data <- dcm.new.data %>% 
     mutate(Vol = ifelse(type=="Blk",3.5,
                         ifelse(type=="Poo", 3.5, Vol)))
dcm.new.data <- dcm.new.data %>% 
     mutate(Area.norm = ifelse(
          test = grepl(paste(IS.dcm.list, collapse="|"), Compound.Name),
                               yes = Area/Vol,
                               no = Area.BlkSub/Vol)) %>%
      mutate(Area.norm = ifelse(Area.norm<0, NA, Area.norm))
```

##remove 21-24 since they don't have the right extraction standards
```{r eval=F}
dcm.new.data.2 <- dcm.new.data %>%
     filter(SampID!="21",SampID!="22",SampID!="23",SampID!="24")
```

##Calculate mean values for each IS
```{r ISmeans}
IS.aq.dat <- new.data[grepl(paste(IS.aq.list, collapse="|"),
                            new.data$Compound.Name),]
IS.aq.dat <- IS.aq.dat %>% 
     select(Compound.Name, Area.norm, Replicate.Name) %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     filter(!grepl("Poo",Replicate.Name)) 
IS.aq.means <- IS.aq.dat %>% 
     filter(!grepl("Blk", Replicate.Name)) %>%
     group_by(Compound.Name) %>%
     summarize(ave = mean(Area.norm))
```

```{r dcm.ISmeans}
IS.dcm.dat <- dcm.new.data.2[grepl(paste(IS.dcm.list, collapse="|"),
                            dcm.new.data.2$Compound.Name),]
IS.dcm.dat <- IS.dcm.dat %>% 
     select(Compound.Name, Area.norm, Replicate.Name) %>%
     filter(!grepl("Blk",Replicate.Name)) %>%
     filter(!grepl("Poo",Replicate.Name)) 
IS.dcm.means <- IS.dcm.dat %>% 
     filter(!grepl("Blk", Replicate.Name)) %>%
     group_by(Compound.Name) %>%
     summarize(ave = mean(Area.norm))
```

##Normalize to each internal Standard for which there aren't NAs
Actually scaling/adjusting by dividing by IS peak area and multiplying by the average IS peak area
```{r aqNormalization, echo=FALSE}
##Clean up data - remove rows that only have Area in the blank samples
wArea.aq <- new.data %>%
     select(Replicate.Name,Compound.Name,Area.norm) %>%
     spread(key=Compound.Name, value=Area.norm) %>% 
     filter(!grepl("Blk",Replicate.Name)) %>%
     Filter(function(x) !all(is.na(x)), .) %>%
     as.data.frame

## Normalize to ISs
this.IS <- IS.aq.means[[1,1]]
area.norm.aq <- wArea.aq[,-1] %>% 
     sapply(FUN = function(x) x/wArea.aq[,grep(this.IS,
                                               names(wArea.aq))]) %>%
     as_data_frame %>% 
     mutate(Replicate.Name = wArea.aq$Replicate.Name) %>%
     gather(Compound,Area_Norm, -Replicate.Name)
this.mean <- IS.aq.means %>% filter(Compound.Name==this.IS) %>% 
     select(ave) %>% as.numeric
area.norm.aq <- area.norm.aq %>% mutate(Area_Norm = Area_Norm*this.mean)
key <- ncol(area.norm.aq)
count <- length(which(!is.na(area.norm.aq$Area_Norm)))/
               length(unique(area.norm.aq$Replicate.Name))
names(area.norm.aq)[key] <- paste(this.IS,"Norm.Area",sep=".")
print(paste(1, this.IS, count, sep="-"))

for (i in 2:length(IS.aq.list)){
     if(!is.na(IS.aq.means[[i,2]])){
     this.IS <- IS.aq.means[[i,1]]
     if(length(wArea.aq[,grep(this.IS, names(wArea.aq))])!=0){
          this.norm <- wArea.aq[,-1] %>% 
               sapply(FUN = function(x) x/wArea.aq[,grep(this.IS,
                                                  names(wArea.aq))]) %>%
               as_data_frame %>% 
               mutate(Replicate.Name = wArea.aq$Replicate.Name) %>%
               gather(Compound,Area_Norm, -Replicate.Name)
          this.mean <- IS.aq.means %>% filter(Compound.Name==this.IS) %>% 
               select(ave) %>% as.numeric
          this.norm <- this.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
          key <- ncol(area.norm.aq)
          area.norm.aq[,key+1] <- this.norm$Area_Norm
          names(area.norm.aq)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          print(paste(i, this.IS, count, sep="-"))
     }
     }
}
```

```{r dcmNormalization, echo=FALSE}
##Clean up data - remove rows that only have Area in the blank samples
wArea.dcm <- dcm.new.data.2 %>%
     select(Replicate.Name,Compound.Name,Area.norm) %>%
     spread(key=Compound.Name, value=Area.norm) %>% 
     filter(!grepl("Blk",Replicate.Name)) %>%
     Filter(function(x) !all(is.na(x)), .) %>%
     as.data.frame

## Normalize to ISs
this.IS <- IS.dcm.means[[1,1]]
area.norm.dcm <- wArea.dcm[,-1] %>% 
     sapply(FUN = function(x) x/wArea.dcm[,grep(this.IS,
                                               names(wArea.dcm))]) %>%
     as_data_frame %>% 
     mutate(Replicate.Name = wArea.dcm$Replicate.Name) %>%
     gather(Compound,Area_Norm, -Replicate.Name)
this.mean <- IS.dcm.means %>% filter(Compound.Name==this.IS) %>% 
     select(ave) %>% as.numeric
area.norm.dcm <- area.norm.dcm %>% mutate(Area_Norm = Area_Norm*this.mean)
key <- ncol(area.norm.dcm)
count <- length(which(!is.na(area.norm.dcm$Area_Norm)))/
               length(unique(area.norm.dcm$Replicate.Name))
names(area.norm.dcm)[key] <- paste(this.IS,"Norm.Area",sep=".")
print(paste(1, this.IS, count, sep="-"))

for (i in 2:length(IS.dcm.list)){
     if(!is.na(IS.dcm.means[[i,2]])){
     this.IS <- IS.dcm.means[[i,1]]
     if(length(wArea.dcm[,grep(this.IS, names(wArea.dcm))])!=0){
          this.norm <- wArea.dcm[,-1] %>% 
               sapply(FUN = function(x) x/wArea.dcm[,grep(this.IS,
                                                  names(wArea.dcm))]) %>%
               as_data_frame %>% 
               mutate(Replicate.Name = wArea.dcm$Replicate.Name) %>%
               gather(Compound,Area_Norm, -Replicate.Name)
          this.mean <- IS.dcm.means %>% filter(Compound.Name==this.IS) %>% 
               select(ave) %>% as.numeric
          this.norm <- this.norm %>% mutate(Area_Norm = Area_Norm*this.mean)
          key <- ncol(area.norm.dcm)
          area.norm.dcm[,key+1] <- this.norm$Area_Norm
          names(area.norm.dcm)[key+1] <- paste(this.IS,"Norm.Area",sep=".")
          count <- length(which(!is.na(this.norm$Area_Norm)))/
               length(unique(this.norm$Replicate.Name))
          print(paste(i, this.IS, count, sep="-"))
     }
     }
}
```

#re-join
```{r echo=FALSE}
area.norm.aq <- area.norm.aq %>% 
     mutate(Run.Cmpd = paste(Replicate.Name, Compound))
new.data <- new.data %>% 
     mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))
aq.dat <- full_join(new.data, area.norm.aq)
aq.dat <- aq.dat %>% select(-Compound)
glimpse(aq.dat)
```

```{r dcm, echo=FALSE}
area.norm.dcm <- area.norm.dcm %>% 
     mutate(Run.Cmpd = paste(Replicate.Name, Compound))
dcm.new.data.2 <- dcm.new.data.2 %>% 
     mutate(Run.Cmpd = paste(Replicate.Name, Compound.Name))
dcm.dat <- full_join(dcm.new.data.2, area.norm.dcm)
dcm.dat <- dcm.dat %>% select(-Compound)
glimpse(dcm.dat)
```

##Compare normalizations
get rsd for each normalization of each compound for each sample
```{r aq, echo=FALSE}
no.blank.dat.aq <- aq.dat %>% filter(type =="Smp" | type =="Poo") 
rsd.stats.aq <- no.blank.dat.aq %>% select(-Replicate.Name,
                                           -Retention.Time,
                                           -Notes, -Height, -IR, -S.N,
                                           -Run.Cmpd,
                                           -runDate, -replicate,
                                           -Vol, -Date, -Area,
                                           -BlkRatio,
                                           -Area.BlkSub,
                                           -`d-IAA.Norm.Area`,
                                           -newArea) %>%
     gather(Normer, Value, -Compound.Name, -type, - SampID) %>%
     group_by(Compound.Name, type, SampID, Normer) %>%
     summarize(m = mean(Value, na.rm=T), 
                sd = sd(Value, na.rm=T), rsd = sd/m) %>%
     filter(!is.na(m))
```
get rsd for each normalization of each compound for each sample
```{r dcm, echo=FALSE}
no.blank.dat.dcm <- dcm.dat %>% filter(type =="Smp" | type =="Poo") 
rsd.stats.dcm <- no.blank.dat.dcm %>% select(-Replicate.Name,
                                           -Retention.Time,
                                           -Notes, -Height, -IR, -S.N,
                                           -Run.Cmpd,
                                           -runDate, -replicate,
                                           -Vol, -Date, -Area,
                                           -BlkRatio,
                                           -Area.BlkSub,
                                           # -`D8-Arachidonic Acid.Norm.Area`,
                                           -newArea) %>%
     gather(Normer, Value, -Compound.Name, -type, - SampID) %>%
     group_by(Compound.Name, type, SampID, Normer) %>%
     summarize(m = mean(Value, na.rm=T), 
                sd = sd(Value, na.rm=T), rsd = sd/m) %>%
     filter(!is.na(m))
```

##Cleanup RSD data and add a minimizing column 
#==> dynamic model of normalization
```{r newModeMaker}
ModelMaker <- function(rsd.stats){
     rsd.clean <- rsd.stats %>% filter(!is.na(rsd)) %>%
          filter(Normer!="Area.norm")
     SmpModel <- rsd.clean %>% filter(type=="Smp") %>%
          select(-m, -sd) %>%
          group_by(Compound.Name, Normer) %>%
          summarize(Mean = mean(rsd, na.rm=T)) %>%
          summarize(SmpModelRSD = min(Mean),
                    Smp.Picked.IS = unique(Normer)[which.min(Mean)])
     PooModel <- rsd.clean %>% filter(type=="Poo") %>%
          select(-m, -sd) %>%
          group_by(Compound.Name, Normer) %>%
          summarize(Mean = mean(rsd, na.rm=T)) %>%
          summarize(PooModelRSD = min(Mean),
                    Poo.Picked.IS = unique(Normer)[which.min(Mean)])
     Models <- full_join(PooModel, SmpModel)
     Models
}
```

```{r}
models.aq <- ModelMaker(rsd.stats.aq)
models.dcm <- ModelMaker(rsd.stats.dcm)
```

##Merge Models and RSD data all together
```{r echo=FALSE}
rsd.total.aq <- full_join(rsd.stats.aq,models.aq)
rsd.total.dcm <- full_join(rsd.stats.dcm, models.dcm)
```

##Adjust the pooled model 
so that it normalizes metabolites that we have internal standards for to 
the appropriate internal standard. Otherwise keeps the pooled model. Names this new model "PooPlus"
```{r aq}
rsd.total.aq <- rsd.total.aq %>%
     mutate(PooPlus.IS = Poo.Picked.IS) %>%
     mutate(PooPlusModelRSD = PooModelRSD)
used.IS <- IS.aq.means %>% filter(!is.na(ave)) %>% select(Compound.Name)
used.IS <- used.IS[[1]][2:3]     

for (i in 1:nrow(rsd.total.aq)){
     cmpd <- rsd.total.aq$Compound.Name[i]
     if(length(grep(cmpd, used.IS))>0){
          newIS <- paste0(IS.list[grep(cmpd, IS.list)],".Norm.Area")
          rsd.total.aq$PooPlus.IS[i] <- newIS
          rsd.total.aq$PooPlusModelRSD[i] <- mean(filter(rsd.total.aq, 
                                                 Compound.Name == cmpd, 
                                                 Normer == newIS, 
                                                 type == "Poo")$rsd)
     }
     ## if the poo didn't pick a normalizing IS, use the samp picked one
     if(is.na(rsd.total.aq$Poo.Picked.IS[i]) &
        !is.na(rsd.total.aq$Smp.Picked.IS[i])){
          rsd.total.aq$PooPlus.IS[i] <- rsd.total.aq$Smp.Picked.IS[i]
          rsd.total.aq$PooPlusModelRSD[i] <-
               rsd.total.aq$SmpModelRSD[i]
     }
}
```

```{r dcm}
rsd.total.dcm <- rsd.total.dcm %>%
     mutate(PooPlus.IS = Poo.Picked.IS) %>%
     mutate(PooPlusModelRSD = PooModelRSD)
used.IS <- IS.dcm.means %>% filter(!is.na(ave)) %>% select(Compound.Name)
used.IS <- used.IS[[1]][c(1,3)]     

for (i in 1:nrow(rsd.total.dcm)){
     cmpd <- rsd.total.dcm$Compound.Name[i]
     if(length(grep(cmpd, used.IS))>0){
          newIS <- paste0(IS.list[grep(cmpd, IS.list)],".Norm.Area")
          rsd.total.dcm$PooPlus.IS[i] <- newIS
          rsd.total.dcm$PooPlusModelRSD[i] <- mean(filter(rsd.total.dcm, 
                                                 Compound.Name == cmpd, 
                                                 Normer == newIS, 
                                                 type == "Poo")$rsd)
     }
     ## if the poo didn't pick a normalizing IS, use the samp picked one
     if(is.na(rsd.total.dcm$Poo.Picked.IS[i]) &
        !is.na(rsd.total.dcm$Smp.Picked.IS[i])){
          rsd.total.dcm$PooPlus.IS[i] <- rsd.total.dcm$Smp.Picked.IS[i]
          rsd.total.dcm$PooPlusModelRSD[i] <-
               rsd.total.dcm$SmpModelRSD[i]
     }
}
```
##Get RSD data I can plot
```{r aq, echo=F}
models.aq <- rsd.total.aq %>% ungroup %>%
     select(Compound.Name,
            Poo.Picked.IS,Smp.Picked.IS,PooPlus.IS) %>%
     filter(!is.na(PooPlus.IS)) %>%
     group_by(Compound.Name) %>%
     summarize(PooModel.IS = unique(Poo.Picked.IS),
               SmpModel.IS = unique(Smp.Picked.IS),
               PooPlusModel.IS = unique(PooPlus.IS)) %>%
     mutate(SmpModel.IS = ifelse(SmpModel.IS=="", 
                                 yes = NA,
                                 no = SmpModel.IS))

rsd.plot.data <- rsd.total.aq %>% select(-m,-sd,-PooModelRSD,-SmpModelRSD,-PooPlusModelRSD, -Poo.Picked.IS, -Smp.Picked.IS, -PooPlus.IS) %>% spread(Normer, rsd)

#For if there are only the two IS---------
rsd.plot.data2 <- full_join(models.aq, rsd.plot.data)
rsd.plot.data2 <- rsd.plot.data2 %>%
     mutate(PooPlusRSD =
                 ifelse(PooPlusModel.IS=="D-Phenylalanine.Norm.Area",
                                `D-Phenylalanine.Norm.Area`,
                                ifelse(PooPlusModel.IS==
                                            "Vitamin B2_IS.Norm.Area",
                                       `Vitamin B2_IS.Norm.Area`,NA)))

# ##For if there are all the IS----------
# rsd.plot.data2 <- full_join(models.aq, rsd.plot.data) 
# rsd.plot.data2 <- rsd.plot.data2 %>%
#      mutate(PooPlusRSD =
#                  ifelse(PooPlusModel.IS=="D-Phenylalanine.Norm.Area",
#                                 `D-Phenylalanine.Norm.Area`,
#                                 ifelse(PooPlusModel.IS==
#                                             "Vitamin B2_IS.Norm.Area",
#                                        yes=`Vitamin B2_IS.Norm.Area`,
#                                       no=ifelse(PooPlusModel.IS==
#                                             "d4-Tryptamine.Norm.Area",
#                                        yes=`d4-Tryptamine.Norm.Area`,
#                                        no=ifelse(PooPlusModel.IS==
#                                             "d-IAA.Norm.Area",
#                                        yes=`d-IAA.Norm.Area`,
#                                        no=ifelse(PooPlusModel.IS==
#                                             "D-Tryptophan.Norm.Area",
#                                        yes=`D-Tryptophan.Norm.Area`,
#                                        no = NA))))))

rsd.plot.data.long <- rsd.plot.data2 %>% 
     select(-PooModel.IS, -SmpModel.IS, -PooPlusModel.IS) %>%
     gather(Normer, RSD, -Compound.Name, -type, -SampID)
```

```{r dcm, echo=F}
models.dcm <- rsd.total.dcm %>% ungroup %>%
     select(Compound.Name,
            Poo.Picked.IS,Smp.Picked.IS,PooPlus.IS) %>%
     filter(!is.na(PooPlus.IS)) %>%
     group_by(Compound.Name) %>%
     summarize(PooModel.IS = unique(Poo.Picked.IS),
               SmpModel.IS = unique(Smp.Picked.IS),
               PooPlusModel.IS = unique(PooPlus.IS)) %>%
     mutate(SmpModel.IS = ifelse(SmpModel.IS=="", 
                                 yes = NA,
                                 no = SmpModel.IS))

rsd.plot.data.dcm <- rsd.total.dcm %>% select(-m,-sd,-PooModelRSD,-SmpModelRSD,-PooPlusModelRSD, -Poo.Picked.IS, -Smp.Picked.IS, -PooPlus.IS) %>% spread(Normer, rsd)

#For if there are only the two IS---------
# rsd.plot.data.dcm2 <- full_join(models.dcm, rsd.plot.data.dcm)
# rsd.plot.data.dcm2 <- rsd.plot.data.dcm2 %>%
#      mutate(PooPlusRSD =
#                  ifelse(PooPlusModel.IS=="D-Phenylalanine.Norm.Area",
#                                 `D-Phenylalanine.Norm.Area`,
#                                 ifelse(PooPlusModel.IS==
#                                             "Vitamin B2_IS.Norm.Area",
#                                        `Vitamin B2_IS.Norm.Area`,NA)))

# For if there are all the IS----------
rsd.plot.data.dcm2 <- full_join(models.dcm, rsd.plot.data.dcm)
rsd.plot.data.dcm2 <- rsd.plot.data.dcm2 %>%
     mutate(PooPlusRSD =
                 ifelse(PooPlusModel.IS=="D-Phenylalanine.Norm.Area",
                                `D-Phenylalanine.Norm.Area`,
                                ifelse(PooPlusModel.IS==
                                            "Vitamin B2_IS.Norm.Area",
                                       yes=`Vitamin B2_IS.Norm.Area`,
                                      no=ifelse(PooPlusModel.IS==
                                            "D8-Arachidonic Acid.Norm.Area",
                           yes=`D8-Arachidonic Acid.Norm.Area`,no = NA))))

rsd.plot.data.dcm.long <- rsd.plot.data.dcm2 %>% 
     select(-PooModel.IS, -SmpModel.IS, -PooPlusModel.IS) %>%
     gather(Normer, RSD, -Compound.Name, -type, -SampID)
```

#plot rsd comparisions
```{r rsdPlots}
ggplot(rsd.plot.data.dcm.long, aes(x= Normer, y = RSD)) + geom_boxplot() +
     ylim(0,.5)
```

##Write rsd data
```{r eval=FALSE}
write.csv(rsd.plot.data2, "Cyano_Aq_diel_1_normalization_rsd_allDat_blkSub_backfill_new0926.csv")
write.csv(rsd.plot.data.dcm2, "Cyano_dcm_diel_1_normalization_rsd_allNorms.no21to24_blkSub_backfill_new0926.csv")
```

##Get Actual Normalized data not just the rsds
```{r aq, echo=FALSE}
aq.dat <- aq.dat %>% filter(!is.na(Replicate.Name))
dat.join.aq <- full_join(aq.dat, models.aq)

PooModel <- c()
SmpModel <-c()
PooPlusModel <- c()
for (i in 1:nrow(dat.join.aq)){
     if(!is.na(dat.join.aq$PooModel.IS[i])){
          PooModel <- c(PooModel,dat.join.aq[i,dat.join.aq$PooModel.IS[i]])
     } else {
          PooModel <- c(PooModel,NA)
     }
     if(!is.na(dat.join.aq$SmpModel.IS[i])){
          SmpModel <- c(SmpModel,dat.join.aq[i,
                                            dat.join.aq$SmpModel.IS[i]])
     } else {
          SmpModel <- c(SmpModel,NA)
     }
     if(!is.na(dat.join.aq$PooPlusModel.IS[i])){
          PooPlusModel <- c(PooPlusModel,
                            dat.join.aq[i,dat.join.aq$PooPlusModel.IS[i]])
     } else {
          PooPlusModel <- c(PooPlusModel,NA)
     }
}
dat.join.aq$PooModel <- PooModel
dat.join.aq$SmpModel <- SmpModel
dat.join.aq$PooPlusModel <- PooPlusModel

dat.join.aq <- dat.join.aq %>% unnest(PooModel) %>% unnest(SmpModel) %>%
     unnest(PooPlusModel)
```

```{r dcm, echo=FALSE}
dcm.dat <- dcm.dat %>% filter(!is.na(Replicate.Name))
dat.join.dcm <- full_join(dcm.dat, models.dcm)

PooModel <- c()
SmpModel <-c()
PooPlusModel <- c()
for (i in 1:nrow(dat.join.dcm)){
     if(!is.na(dat.join.dcm$PooModel.IS[i])){
          PooModel <- c(PooModel,dat.join.dcm[i,dat.join.dcm$PooModel.IS[i]])
     } else {
          PooModel <- c(PooModel,NA)
     }
     if(!is.na(dat.join.dcm$SmpModel.IS[i])){
          SmpModel <- c(SmpModel,dat.join.dcm[i,
                                            dat.join.dcm$SmpModel.IS[i]])
     } else {
          SmpModel <- c(SmpModel,NA)
     }
     if(!is.na(dat.join.dcm$PooPlusModel.IS[i])){
          PooPlusModel <- c(PooPlusModel,
                            dat.join.dcm[i,dat.join.dcm$PooPlusModel.IS[i]])
     } else {
          PooPlusModel <- c(PooPlusModel,NA)
     }
}
dat.join.dcm$PooModel <- PooModel
dat.join.dcm$SmpModel <- SmpModel
dat.join.dcm$PooPlusModel <- PooPlusModel

dat.join.dcm <- dat.join.dcm %>% unnest(PooModel) %>% unnest(SmpModel) %>%
     unnest(PooPlusModel) %>% filter(!is.na(SampID))
```
##Write out that full normalized data
```{r}
write.csv(dat.join.aq, "Cyano_Aq_diel_1_Samples_Normalized_BlkSub_backfill_2IS_new0926.csv")
dat.join.aq.noblk <- dat.join.aq %>% 
     filter(type!="Blk")
write.csv(dat.join.aq.noblk, "Cyano_Aq_diel_1_Samples_Normalized_BlkSub_backfill_2IS_noblk_new0926.csv")

write.csv(dat.join.dcm, "Cyano_DCM_diel_1_Samples_Normalized_BlkSub_backfill_allIS_new0926.csv")
dat.join.dcm.noblk <- dat.join.dcm %>% filter(type!="Blk")
write.csv(dat.join.dcm.noblk, "Cyano_DCM_diel_1_Samples_Normalized_BlkSub_backfill_allIS_noblk_new0926.csv")
```
